High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-23 doc/asplos2013/impl.tex
move discussion of lock payloads to interface section

Locks are often used to mediate access to data, and the size of that data is often small.  To
support this common case we allow locks to be created with an associated
%avoid adding the latency of inter-node data copies to critical sections, a lock may also be created with a
{\em payload}, a small (i.e. less than 4KB - the limit for an active message payload) 
piece of data that is moved around with the lock and guaranteed to be coherent while 
the lock is held.

@@

When a lock request is made, an event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiving node is no
longer the owner, it will forward the message on to the next node known
to be the owner of the lock.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message.  A lock transfer message includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.

>>>>

@@

When a lock request is made, an event is created to track when the grant occurs.  The current node then
examines its copy of the lock data structure to determine if it is the owner.
%(creating it if this the first operation for that lock on that node).  
If the current node is the owner and the lock isn't
held, the lock is granted immediately and the event is triggered.  If the lock is held, the event is added
to the list of local waiters.  Note that the event associated with
the lock request is the only data that must be stored.
%(This is the only thing that is stored.  The actual ``requestor'' of the lock
%is implicitly captured in which operations are dependent on the lock grant event.)  
If the current node isn't the owner, a {\em lock request} active message is
sent to the most recently known owner.  If the receiving node is no
longer the owner, it will forward the message on to the next node known
to be the owner of the lock.
%(If that information proves to be stale, the request is forwarded
%by the receiving node, eventually catching up to the current owner.)  
If the current owner's status shows
the lock is currently held, the bit for the requesting node is set in remote waiters bitmask.  If the lock is
not held, the ownership of the lock is given to the requesting node via a {\em lock transfer} active
message.  A lock transfer message includes the bitmask of remote waiters and an up-to-date copy 
of the lock's payload.  (The inclusion of the payload in the active message is the reason for the 
4KB limit on a payload's size.)

### 2012-07-23 doc/asplos2013/interface.tex
move discussion of lock payloads to interface section

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is shown in Listing~\ref{lst:machineapi} and allows the
initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (line 15-16).
Any running task can get a pointer to the {\tt Machine} object with the
{\tt get\_machine} method (line 18) and use it to obtain lists of all the {\tt Memory} (line 19) or 
{\tt Processor} (line 20) handles in the system.  Using the affinity methods (lines 22-25),
the application can also determine:
\begin{itemize} \itemsep1pt \parskip0pt \parsep0pt
\item for each {\tt Processor}, which {\tt Memory} objects are accessible and with what performance, and
\item which pairs of {\tt Memory} objects are able to efficiently copy data between them.
\end{itemize}

>>>>

When a lock is used to mediate access to a piece of data, the first thing a task will usually
do after being granted the lock is access the data.  To be able to hide the latency of that
access as well, an additional feature of locks in our interface is the association of a lock with a
{\em payload}, a small (i.e. less than 4KB) 
piece of data that is moved around with the lock and guaranteed to be coherent while 
the lock is held.  The size of the payload is specified when the lock is created (line 7) and
the pointer to the local copy of the payload can be obtained from the {\tt payload\_ptr} method
(line 8).

Locks are often used to mediate access to data, and the size of that data is often small.  To
support this common case we allow locks to be created with an associated
%avoid adding the latency of inter-node data copies to critical sections, a lock may also be created with a
{\em payload}, a small (i.e. less than 4KB - the limit for an active message payload) 
piece of data that is moved around with the lock and guaranteed to be coherent while 
the lock is held.

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is shown in Listing~\ref{lst:machineapi} and allows the
initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (line 15-16).
Any running task can get a pointer to the {\tt Machine} object with the
{\tt get\_machine} method (line 18) and use it to obtain lists of all the {\tt Memory} (line 19) or 
{\tt Processor} (line 20) handles in the system.  Using the affinity methods (lines 22-25),
the application can also determine which {\tt Memory} objects are accessible by each
{\tt Processor},and with what performance, and which pairs of {\tt Memory} objects are able to
efficiently copy data between them.
